Modal type theory as based on the intuitionistic modal logic IEL$^{-}$

### 2018-05-08 article.tex
another one update

  \begin{minipage}{0.5\textwidth}
    \begin{flushleft}
    \begin{prooftree}
      \AxiomC{$\Gamma, A \vdash B$}
      \RightLabel{$\to_I$}
      \UnaryInfC{$\Gamma \vdash A \to B$}
    \end{prooftree}
    \begin{prooftree}
      \AxiomC{$\Gamma \vdash A$}
      \AxiomC{$\Gamma \vdash B$}
      \RightLabel{$\land_I$}
      \BinaryInfC{$\Gamma \vdash A \land B$}
    \end{prooftree}
  	\begin{prooftree}
    	\AxiomC{$\Gamma \vdash A$}
    	\RightLabel{${\Box_I}_1$}
    	\UnaryInfC{$\Gamma \vdash \Box A$}
  \end{prooftree}
    \end{flushleft}
  \end{minipage}
  \begin{minipage}{0.5\textwidth}
    \begin{flushright}
    \begin{prooftree}
      \AxiomC{$\Gamma \vdash A \to B$}
      \AxiomC{$\Gamma \vdash A$}
      \RightLabel{$\to_E$}
      \BinaryInfC{$\Gamma \vdash B$}
    \end{prooftree}
    \begin{prooftree}
      \AxiomC{$\Gamma \vdash A_1 \land A_2$}
      \RightLabel{$\land_E, i \in {1,2}$}
      \UnaryInfC{$\Gamma \vdash A_i$}
    \end{prooftree}
  	\begin{prooftree}
    \AxiomC{$\Gamma \vdash \Box A_1, \dots, \Gamma \vdash \Box A_n $}
    \AxiomC{$A_1,\dots,A_n \vdash B$}
    \RightLabel{${\Box_I}_2$}
    \BinaryInfC{$\Gamma \vdash \Box B$}
  	\end{prooftree}
    \end{flushright}
  \end{minipage}
  \end{defin}

@@

    \begin{prooftree}
    \AxiomC{ $\Gamma \vdash M : A_1 \times A_2$ }
    \RightLabel{$\times_e$, $i \in \{ 1, 2 \}$}
    \UnaryInfC{$\Gamma \vdash \pi_i M : A_i$}
    \end{prooftree}

@@

\begin{prooftree}
\AxiomC{$x : A \vdash x : A$}
\UnaryInfC{$x : A \vdash {\bf box \:} x : \Box A$}
\UnaryInfC{$\vdash (\lambda x. {\bf box \: } x) : A \to \Box A$}
\end{prooftree}

\begin{prooftree}
\AxiomC{$f : \Box (A \to B) \vdash f : \Box (A \to B)$}
\AxiomC{$x : \Box A \vdash x : \Box A $}
\AxiomC{$g : A \to B \vdash g : A \to B$}
\AxiomC{$y : A \vdash y : A$}
\RightLabel{$\to_e$}
\BinaryInfC{$g : A \to B, y : A \vdash g y : B$}
\RightLabel{$\text{let}_{\Box}$}
\TrinaryInfC{$f : \Box (A \to B), x : \Box A \vdash {\bf let \: box \:} g, y  = f, x {\: \bf in \:} g y : \Box B$}
\UnaryInfC{$f : \Box (A \to B) \vdash \lambda x. {\bf let \: box \:} g, y = f, x {\: \bf in \:} g y : \Box A \to \Box B$}
\UnaryInfC{$\vdash \lambda f. \lambda x. {\bf let \: box \:} g, y = f, x {\: \bf in \:} g y : \Box (A \to B) \to \Box A \to \Box B$}
\end{prooftree}

@@

\begin{enumerate}
  \item $FV(x) = \{ x \}$;
  \item $FV(\lambda x. M) = FV(M) \setminus \{ x\}$;
  \item $FV(M N) = FV(M) \cup FV(N);$
  \item $FV(\langle M,N \rangle) = FV(M) \cup FV(N)$;
  \item $FV(\pi_i M) = FV(M)$, $i \in \{ 1, 2\}$;
  \item $FV(\text{\bf box } M) = FV(M)$;
  \item $FV({\bf let \: box} \: \vec{x} = \vec{M} {\: \bf in \:} N) = \bigcup \limits_{i = 1}^n FV(M), \text{where $n = |\vec{M}|$}$.
\end{enumerate}
  \end{defin}

@@

\begin{enumerate}
  \item $x [x := N] = N$, $x [y := N] = x$;
  \item $(M N) [x := N] = M[x := N] N [x := N]$;
  \item $(\lambda x. M) [x := N] = \lambda x. M [y := N]$, $y \in FV(M)$;
  \item $(M, N)[x := P] = (M[x := P], N [x := P])$;
  \item $(\pi_i M) [x := P] = \pi_i (M[x := P])$, $i \in \{ 1, 2\}$;
  \item $({\bf box \: } M) [x := P] = {\bf box \: } (M [x := P])$;
  \item $({\bf let \: box \:}\vec{x} = \vec{M} {\: \bf in \:} N) [y := P] = {\bf let \: box \:} \vec{x} = (\vec{M} [y := P]) {\: \bf in \:} N$.
\end{enumerate}
  \end{defin}

@@

\begin{enumerate}
  \item $(\lambda x. M) N \rightarrow_{\beta} M [x := N]$;
  \item $\pi_1 \langle M, N \rangle \rightarrow_{\beta} M$;
  \item $\pi_2 \langle M, N \rangle \rightarrow_{\beta} N$;
  \item ${\bf let \: box \:} \vec{x}, y, \vec{z} = \vec{M}, {\bf let \: box \: } \vec{w} = \vec{N} {\: \bf in \: } Q, \vec{P} {\: in \:} R \rightarrow_{\beta} \\
  {\bf let \: box \:} \vec{x}, \vec{w}, \vec{z} = \vec{M}, \vec{N}, \vec{P} {\: \bf in \: } R [y := Q]$
  \item ${\bf let \: box \:} \vec{x} = {\bf box \:} \vec{M} {\: \bf in \:} N \rightarrow_{\beta} {\bf box \:} N [\vec{x} := \vec{M}]$
  \item ${\bf let \: box \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} M \rightarrow_{\beta} {\bf box \:} M$, where \underline{\quad} is an empty sequence of terms.
  \item $\lambda x. f x \rightarrow_{\eta} f$;
  \item $\langle \pi_1 P, \pi_2 P \rangle \rightarrow_{\eta} P$;
  \item ${\bf let \: box \:} x = M {\: \bf in \: } x \rightarrow_{\eta} M$;
\end{enumerate}
  \end{defin}

@@

Now we will prove standard lemmas for contexts in type systems\footnote{We will not prove cases with
$\to$-constructor, they are proved standardly in the same lemmas for simply typed lambda calculus, for
example see \cite{Neder} \cite{Morten}. We will consider only modal cases}:

@@

\begin{enumerate}
  \item If $\Gamma \vdash M : A$ and $\Gamma \subseteq \Delta$, then $\Delta \vdash M : A$;
  \item If $\Gamma \vdash M : A$, then $\Delta \vdash M : A$, where $\Delta = \{ x_i : A_i \: | \: (x_i : A_i) \in \Gamma \: \& \: x_i \in FV(M) \}$
  \item If $\Gamma, x : A \vdash M : B$ and $\Gamma \vdash N : A$, then $\Gamma \vdash M [x := N] : B$.
\end{enumerate}
\end{lemma}

@@

Induction on the derivation of $\Gamma \vdash M : A$.

@@

  If $\Gamma \vdash M : A$ and $M \twoheadrightarrow_r N$, then $\Gamma \vdash N : A$

@@

For cases with axiom, application, abstraction and pairs see \cite{Morten} \cite{Pierce}.

@@

  Let us define the transformation from $\lambda_{\text{IEL}^{-}}$ into the simple type theory with
  $\to$, $\times$ and natural number type $\mathbb{N}$ with additional typing and reduction rules
  \footnote{Strong normalization for stronger system was shown here \cite{Girard}} :

  \begin{prooftree}
  \AxiomC{$ $}
  \UnaryInfC{$\Gamma \vdash 0 : \mathbb{N}$}
  \end{prooftree}

  \begin{prooftree}
  \AxiomC{$\Gamma \vdash n : \mathbb{N}$}
  \UnaryInfC{$\Gamma \vdash {\bf succ \:} n : \mathbb{N}$}
  \end{prooftree}

  \begin{prooftree}
  \AxiomC{$\Gamma \vdash n : \mathbb{N}$}
  \AxiomC{$\Gamma \vdash m : \mathbb{N}$}
  \BinaryInfC{$\Gamma \vdash n + m : \mathbb{N}$}
  \end{prooftree}

@@

  \begin{defin} Interpretation of terms:
    \begin{enumerate}
      \item $x \in \mathbb{V} \Rightarrow |x| = x$;
      \item $|\lambda x. M| = \lambda x. |M|$;
      \item $|(M N)| = |M| |N|$;
      \item $|\langle M, N \rangle| = \langle |M|, |N| \rangle$;
      \item $|\pi_i M| = \pi_i |M|$, $i \in \{ 1, 2\}$;
      \item $|{\bf box \:} M| = \langle 0, |M| \rangle$;
      \item $|{\bf let \: box \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} M| = \langle 0, |M| \rangle$;
      \item $|{\bf let \: box \:} x = N {\: \bf in \:} M| = \langle \pi_1 |N| , |M| [x := \pi_2 |N|] \rangle$
      \item $|{\bf let \: box \:} \vec{x} = \vec{N} {\: \bf in \:} M| = \langle \sum \limits_{i = 1}^n \pi_1 |N|, |M| [\vec{x} := \pi_2 |\vec{N}|] \rangle$
    \end{enumerate}
  \end{defin}

@@

  Monoidal category is a category $\mathcal{C}$ with additional monoidal structure:
  \begin{enumerate}
    \item A bifunctor $\otimes : \mathcal{C} \times \mathcal{C} \to C$ called tensor product;
    \item Identity object $\mathds{1}$;
    \item A natural isomorphism called associator: $\alpha_{A,B,C}: (A \otimes B) \otimes C \cong A \otimes (B \otimes C)$;
    \item A natural isomorphism called left unitor: $L_A : \mathds{1} \otimes A \cong A$;
    \item A natural isomorphism called right unitor $R_A : A \otimes \mathds{1} \cong A$;
    \item A coherence condition called MacLane pentagon, i.e. the following diagram commutes:

@@

  A monoidal functor $\mathcal{F} : \langle \mathcal{C}, \otimes_1, \mathds{1} \rangle \to \langle \mathcal{D}, \otimes_2, \mathds{1}' \rangle$ is a functor
  $\mathcal{F} : \mathcal{C} \to \mathcal{D}$ with additional natural transformations:

@@

\begin{prooftree}
\AxiomC{$[\![\Gamma \vdash M : A \to B]\!] = [\![M]\!] : [\![\Gamma]\!] \rightarrow [\![B]\!]^{[\![A]\!]}$}
\AxiomC{$[\![\Gamma \vdash N : A]\!] = [\![N]\!] : [\![\Gamma]\!] \rightarrow [\![A]\!]$}
\BinaryInfC{$[\![\Gamma \vdash (M N) : B]\!] = [\![\Gamma]\!] \xrightarrow{\langle [\![M]\!], [\![N]\!]
\rangle} [\![B]\!]^{[\![A]\!]} \times [\![A]\!] \xrightarrow{\epsilon} [\![B]\!] $}
\end{prooftree}

@@

\begin{prooftree}
\AxiomC{$[\![\Gamma \vdash M : A_1 \times A_2]\!] = [\![M]\!] : [\![\Gamma]\!] \rightarrow [\![A_1]\!] \times
[\![A_2]\!]$}
\RightLabel{$i \in \{1,2\}$}
\UnaryInfC{$[\![\Gamma \vdash \pi_i M : A_i]\!] = [\![\Gamma]\!] \xrightarrow{[\![M]\!]} [\![A_1]\!] \times
[\![A_2]\!] \xrightarrow{\pi_i} [\![A_i]\!]$}
\end{prooftree}

@@

\begin{defin} Category $\mathcal{C}(\lambda)$:
\begin{enumerate}
  \item $Ob_{\mathcal{C}} = \{ \hat{A} \: | \: A \in \mathbb{T} \} \cup \{ \mathds{1} \}$;
  \item $Hom_{\mathcal{C}(\lambda)}(\hat{A},\hat{B}) = \{ [x, M] \: | \: x : A \vdash_{\lambda_{\text{IEL}^{-}}} M : B \}$;
  \item Let $[x, M] \in Hom_{\mathcal{C}(\lambda)}(\hat{A},\hat{B})$ and $[y,N] \in Hom_{\mathcal{C}(\lambda)}(\hat{B},\hat{C})$, then $[y,M] \circ [x, M] = [x, N [y := M]]$;
  \item Identity morphism $id_{\hat{A}} = [x,x] \in Hom_{\mathcal{C}(\lambda)(\hat{A}, \hat{A})}$;
  \item $\mathds{1}$ is a terminal object;
  \item $\widehat{A \times B} = \hat{A} \times \hat{B}$;
  \item Canonical projection is defined as $[x, \pi_i x] \in Hom_{\mathcal{C}(\lambda)}(\hat{A_1} \times \hat{A_2},\hat{A_i})$ for $i \in \{ 1, 2 \}$;
  \item $\widehat{A \to B} = \hat{B}^{\hat{A}}$;
  \item Evaluation arrow $\epsilon_{\hat{A},\hat{B}} = [x, (\pi_2 x) (\pi_1 x)] \in Hom_{\mathcal{C}(\lambda)}(\hat{B}^{\hat{A}} \times \hat{A}, \hat{B})$.
\end{enumerate}
\end{defin}

@@

\begin{enumerate}
  \item $\boxdot(g \circ f) = \boxdot g \circ \boxdot f$;
  \item $\boxdot (id_{\hat{A}}) = id_{\boxdot}\hat{A}$.
\end{enumerate}
\end{lemma}

@@

\begin{enumerate}
  \item $\eta: Id_{\mathcal{C}(\lambda)} \Rightarrow \boxdot$, s. t. $\forall \hat{A} \in Ob_{\mathcal{C}(\lambda)}$, $\eta_{\hat{A}} = [x, {\bf box \:} x] \in Hom_{\mathcal{C}(\lambda)}(\hat{A}, \boxdot \hat{A})$;
  \item $\hat{A} \varoast \hat{B} : \boxdot \hat{A} \times \boxdot \hat{B} \to \boxdot (\hat{A} \times \hat{B})$, s. t. $\forall \hat{A}, \hat{B} \in Ob_{\mathcal{C}(\lambda)}, \hat{A} \varoast \hat{B} = [p, {\bf let \: box \:} x,y = \pi_1 p, \pi_2 p {\: \bf in \:} \langle x, y \rangle] \in Hom_{\mathcal{C}(\lambda)}(\boxdot \hat{A} \times \boxdot \hat{B}, \boxdot (\hat{A} \times \hat{B}))$.
\end{enumerate}
\end{defin}

@@

\begin{prooftree}
\AxiomC{$p : \Box A \times \Box B \vdash p : \Box A \times \Box B$}
\UnaryInfC{$p : \Box A \times \Box B \vdash \pi_1 p : \Box A$}
\AxiomC{$p : \Box A \times \Box B \vdash p : \Box A \times \Box B$}
\UnaryInfC{$p : \Box A \times \Box B \vdash \pi_2 p : \Box B$}
\AxiomC{$x : A \vdash x : A$}
\AxiomC{$y : B \vdash y : B$}
\BinaryInfC{$x : A, y : B \vdash \langle x, y \rangle : A \times B$}
\TrinaryInfC{$p : \Box A \times \Box B \vdash {\bf let \: box \:} x,y = \pi_1 p, \pi_2 p {\: \bf in \:} \langle x, y \rangle : \Box (A \times B)$}
\end{prooftree}

>>>>

  \begin{minipage}{0.5\textwidth}
    \begin{flushleft}
    \begin{prooftree}
      \AxiomC{$\Gamma, A \vdash B$}
      \RightLabel{$\to_I$}
      \UnaryInfC{$\Gamma \vdash A \to B$}
    \end{prooftree}
    \begin{prooftree}
      \AxiomC{$\Gamma \vdash A$}
      \AxiomC{$\Gamma \vdash B$}
      \RightLabel{$\land_I$}
      \BinaryInfC{$\Gamma \vdash A \land B$}
    \end{prooftree}
  	\begin{prooftree}
    	\AxiomC{$\Gamma \vdash A$}
    	\RightLabel{${\Box_I}_1$}
    	\UnaryInfC{$\Gamma \vdash \Box A$}
  \end{prooftree}
    \end{flushleft}
  \end{minipage}
  \begin{minipage}{0.5\textwidth}
    \begin{flushright}
    \begin{prooftree}
      \AxiomC{$\Gamma \vdash A \to B$}
      \AxiomC{$\Gamma \vdash A$}
      \RightLabel{$\to_E$}
      \BinaryInfC{$\Gamma \vdash B$}
    \end{prooftree}
    \begin{prooftree}
      \AxiomC{$\Gamma \vdash A_1 \land A_2$}
      \RightLabel{$\land_E, i = 1,2$}
      \UnaryInfC{$\Gamma \vdash A_i$}
    \end{prooftree}
  	\begin{prooftree}
    \AxiomC{$\Gamma \vdash \Box A_1, \dots, \Gamma \vdash \Box A_n $}
    \AxiomC{$A_1,\dots,A_n \vdash B$}
    \RightLabel{${\Box_I}_2$}
    \BinaryInfC{$\Gamma \vdash \Box B$}
  	\end{prooftree}
    \end{flushright}
  \end{minipage}
  \end{defin}

@@

    \begin{prooftree}
    \AxiomC{ $\Gamma \vdash M : A_1 \times A_2$ }
    \RightLabel{$\times_e$, $i = 1, 2$}
    \UnaryInfC{$\Gamma \vdash \pi_i M : A_i$}
    \end{prooftree}

@@

\begin{prooftree}
\AxiomC{$x : A \vdash x : A$}
\RightLabel{$\Box_I$}
\UnaryInfC{$x : A \vdash {\bf box \:} x : \Box A$}
\RightLabel{$\to_I$}
\UnaryInfC{$\vdash (\lambda x. {\bf box \: } x) : A \to \Box A$}
\end{prooftree}

\begin{prooftree}
\AxiomC{$f : \Box (A \to B) \vdash f : \Box (A \to B)$}
\AxiomC{$x : \Box A \vdash x : \Box A $}
\AxiomC{$g : A \to B \vdash g : A \to B$}
\AxiomC{$y : A \vdash y : A$}
\RightLabel{$\to_e$}
\BinaryInfC{$g : A \to B, y : A \vdash g y : B$}
\RightLabel{$\text{let}_{\Box}$}
\TrinaryInfC{$f : \Box (A \to B), x : \Box A \vdash {\bf let \: box \:} g, y  = f, x {\: \bf in \:} g y : \Box B$}
\RightLabel{$\to_I$}
\UnaryInfC{$f : \Box (A \to B) \vdash \lambda x. {\bf let \: box \:} g, y = f, x {\: \bf in \:} g y : \Box A \to \Box B$}
\RightLabel{$\to_I$}
\UnaryInfC{$\vdash \lambda f. \lambda x. {\bf let \: box \:} g, y = f, x {\: \bf in \:} g y : \Box (A \to B) \to \Box A \to \Box B$}
\end{prooftree}

@@

\begin{enumerate}
  \item $FV(x) = \{ x \}$;
  \item $FV(\lambda x. M) = FV(M) \setminus \{ x\}$;
  \item $FV(M N) = FV(M) \cup FV(N);$
  \item $FV(\langle M,N \rangle) = FV(M) \cup FV(N)$;
  \item $FV(\pi_i M) = FV(M)$, $i = 1, 2$;
  \item $FV(\text{\bf box } M) = FV(M)$;
  \item $FV({\bf let \: box} \: \vec{x} = \vec{M} {\: \bf in \:} N) = \bigcup \limits_{i = 1}^n FV(M), \text{where $n = |\vec{M}|$}$.
\end{enumerate}
  \end{defin}

@@

\begin{enumerate}
  \item $x [x := N] = N$, $x [y := N] = x$;
  \item $(M N) [x := N] = M[x := N] N [x := N]$;
  \item $(\lambda x. M) [x := N] = \lambda x. M [y := N]$, $y \in FV(M)$;
  \item $(M, N)[x := P] = (M[x := P], N [x := P])$;
  \item $(\pi_i M) [x := P] = \pi_i (M[x := P])$, $i = 1, 2$;
  \item $({\bf box \: } M) [x := P] = {\bf box \: } (M [x := P])$;
  \item $({\bf let \: box \:}\vec{x} = \vec{M} {\: \bf in \:} N) [y := P] = {\bf let \: box \:} \vec{x} = (\vec{M} [y := P]) {\: \bf in \:} N$.
\end{enumerate}
  \end{defin}

@@

\begin{enumerate}
  \item $(\lambda x. M) N \rightarrow_{\beta} M [x := N]$;
  \item $\pi_1 \langle M, N \rangle \rightarrow_{\beta} M$;
  \item $\pi_2 \langle M, N \rangle \rightarrow_{\beta} N$;
  \item ${\bf let \: box \:} \vec{x}, y, \vec{z} = \vec{M}, {\bf let \: box \: } \vec{w} = \vec{N} {\: \bf in \: } Q, \vec{P} {\: in \:} R \rightarrow_{\beta} \\
  {\bf let \: box \:} \vec{x}, \vec{w}, \vec{z} = \vec{M}, \vec{N}, \vec{P} {\: \bf in \: } R [y := Q]$;
  \item ${\bf let \: box \:} \vec{x} = {\bf box \:} \vec{M} {\: \bf in \:} N \rightarrow_{\beta} {\bf box \:} N [\vec{x} := \vec{M}]$;
  \item ${\bf let \: box \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} M \rightarrow_{\beta} {\bf box \:} M$, where \underline{\quad} is an empty sequence of terms;
  \item $\lambda x. f x \rightarrow_{\eta} f$;
  \item $\langle \pi_1 P, \pi_2 P \rangle \rightarrow_{\eta} P$;
  \item ${\bf let \: box \:} x = M {\: \bf in \: } x \rightarrow_{\eta} M$;
\end{enumerate}
  \end{defin}

@@

Now we will prove standard lemmas for contexts in type systems\footnote{Cases with
$\to$-constructor are proven standardly in similar lemmas for simply typed lambda calculus, for
example see \cite{Neder} \cite{Morten} \cite{Pierce}. We will consider only modal cases}:

@@

\begin{enumerate}
  \item If $\Gamma \vdash M : A$ and $\Gamma \subseteq \Delta$, then $\Delta \vdash M : A$;
  \item If $\Gamma \vdash M : A$, then $\Delta \vdash M : A$, where $\Delta = \{ x : A \: | \: (x : A) \in \Gamma \: \& \: x \in FV(M) \}$
  \item If $\Gamma, x : A \vdash M : B$ and $\Gamma \vdash N : A$, then $\Gamma \vdash M [x := N] : B$.
\end{enumerate}
\end{lemma}

@@

1--3) Induction on the derivation of $\Gamma \vdash M : A$.

@@

  If $\Gamma \vdash M : A$ and $M \twoheadrightarrow_r N$, then $\Gamma \vdash N : A$.

@@

@@

  Let us define the transformation from $\lambda_{\text{IEL}^{-}}$ into the simple type theory with
  $\to$, $\times$ and natural number type $\mathbb{N}$ with additional typing and reduction rules\footnote{Strong normalization for stronger system was shown here \cite{Girard}}:

  \begin{prooftree}
  \AxiomC{$ $}
  \RightLabel{{\bf zero}}
  \UnaryInfC{$\Gamma \vdash 0 : \mathbb{N}$}
  \end{prooftree}

\begin{minipage}{0.45\textwidth}
\begin{flushleft}
  \begin{prooftree}
  \AxiomC{$\Gamma \vdash n : \mathbb{N}$}
  \RightLabel{{\bf succ}}
  \UnaryInfC{$\Gamma \vdash {\bf succ \:} n : \mathbb{N}$}
  \end{prooftree}
\end{flushleft}
\end{minipage}
\begin{minipage}{0.45\textwidth}
\begin{flushright}
  \begin{prooftree}
  \AxiomC{$\Gamma \vdash n : \mathbb{N}$}
  \AxiomC{$\Gamma \vdash m : \mathbb{N}$}
  \RightLabel{{\bf add}}
  \BinaryInfC{$\Gamma \vdash n + m : \mathbb{N}$}
  \end{prooftree}
\end{flushright}
\end{minipage}

@@

  \begin{defin} Interpretation of terms:
    \begin{enumerate}
      \item $x \in \mathbb{V} \Rightarrow |x| = x$;
      \item $|\lambda x. M| = \lambda x. |M|$;
      \item $|(M N)| = |M| |N|$;
      \item $|\langle M, N \rangle| = \langle |M|, |N| \rangle$;
      \item $|\pi_i M| = \pi_i |M|$, $i = 1, 2$;
      \item $|{\bf box \:} M| = \langle 0, |M| \rangle$;
      \item $|{\bf let \: box \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} M| = \langle 0, |M| \rangle$;
      \item $|{\bf let \: box \:} x = N {\: \bf in \:} M| = \langle \pi_1 |N| , |M| [x := \pi_2 |N|] \rangle$
      \item $|{\bf let \: box \:} \vec{x} = \vec{N} {\: \bf in \:} M| = \langle \sum \limits_{i = 1}^n \pi_1 |N|, |M| [\vec{x} := \pi_2 |\vec{N}|] \rangle$
    \end{enumerate}
  \end{defin}

@@

  Monoidal category is a category $\mathcal{C}$ with the following additional structure:
  \begin{enumerate}
    \item A bifunctor $\otimes : \mathcal{C} \times \mathcal{C} \to C$ called tensor product;
    \item Identity object $\mathds{1}$;
    \item A natural isomorphism called associator: $\alpha_{A,B,C}: (A \otimes B) \otimes C \cong A \otimes (B \otimes C)$;
    \item A natural isomorphism called left unitor: $L_A : \mathds{1} \otimes A \cong A$;
    \item A natural isomorphism called right unitor $R_A : A \otimes \mathds{1} \cong A$;
    \item A coherence condition called MacLane pentagon, i.e. the following diagram commutes:

@@

  A monoidal functor $\mathcal{F} : \langle \mathcal{C}, \otimes_1, \mathds{1} \rangle \to \langle \mathcal{D}, \otimes_2, \mathds{1}' \rangle$ is a functor
  $\mathcal{F} : \mathcal{C} \to \mathcal{D}$ with natural transformations:

@@

\begin{prooftree}
\AxiomC{$[\![\Gamma \vdash M : A \to B]\!] = [\![M]\!] : [\![\Gamma]\!] \rightarrow [\![B]\!]^{[\![A]\!]}$}
\AxiomC{$[\![\Gamma \vdash N : A]\!] = [\![N]\!] : [\![\Gamma]\!] \rightarrow [\![A]\!]$}
\BinaryInfC{$[\![\Gamma \vdash (M N) : B]\!] = [\![\Gamma]\!] \xrightarrow{\langle [\![M]\!], [\![N]\!]
\rangle} [\![B]\!]^{[\![A]\!]} \times [\![A]\!] \xrightarrow{\epsilon_{[\![A]\!],[\![B]\!]}} [\![B]\!] $}
\end{prooftree}

@@

\begin{prooftree}
\AxiomC{$[\![\Gamma \vdash M : A_1 \times A_2]\!] = [\![M]\!] : [\![\Gamma]\!] \rightarrow [\![A_1]\!] \times
[\![A_2]\!]$}
\RightLabel{$i = 1,2$}
\UnaryInfC{$[\![\Gamma \vdash \pi_i M : A_i]\!] = [\![\Gamma]\!] \xrightarrow{[\![M]\!]} [\![A_1]\!] \times
[\![A_2]\!] \xrightarrow{\pi_i} [\![A_i]\!]$}
\end{prooftree}

@@

\begin{defin} Category $\mathcal{C}(\lambda)$:
\begin{enumerate}
  \item $Ob_{\mathcal{C}} = \{ \hat{A} \: | \: A \in \mathbb{T} \} \cup \{ \mathds{1} \}$;
  \item $Hom_{\mathcal{C}(\lambda)}(\hat{A},\hat{B}) = \{ [x, M] \: | \: x : A \vdash_{\lambda_{\text{IEL}^{-}}} M : B \}$;
  \item Let $[x, M] \in Hom_{\mathcal{C}(\lambda)}(\hat{A},\hat{B})$ and $[y,N] \in Hom_{\mathcal{C}(\lambda)}(\hat{B},\hat{C})$, then $[y,M] \circ [x, M] = [x, N [y := M]]$;
  \item Identity morphism $id_{\hat{A}} = [x,x] \in Hom_{\mathcal{C}(\lambda)(\hat{A}, \hat{A})}$;
  \item $\mathds{1}$ is a terminal object;
  \item $\widehat{A \times B} = \hat{A} \times \hat{B}$;
  \item Canonical projection is defined as $[x, \pi_i x] \in Hom_{\mathcal{C}(\lambda)}(\hat{A_1} \times \hat{A_2},\hat{A_i})$ for $i = 1, 2$;
  \item $\widehat{A \to B} = \hat{B}^{\hat{A}}$;
  \item Evaluation arrow $\epsilon_{\hat{A},\hat{B}} = [x, (\pi_1 x) (\pi_2 x)] \in Hom_{\mathcal{C}(\lambda)}(\hat{B}^{\hat{A}} \times \hat{A}, \hat{B})$.
\end{enumerate}
\end{defin}

@@

\begin{enumerate}
  \item $\boxdot(g \circ f) = \boxdot g \circ \boxdot f$;
  \item $\boxdot (id_{\hat{A}}) = id_{\boxdot \hat{A}}$.
\end{enumerate}
\end{lemma}

@@

\begin{enumerate}
  \item $\eta: Id_{\mathcal{C}(\lambda)} \Rightarrow \boxdot\text{, s. t. }\forall \hat{A} \in Ob_{\mathcal{C}(\lambda)}$, $\eta_{\hat{A}} = [x, {\bf box \:} x] \in Hom_{\mathcal{C}(\lambda)}(\hat{A}, \boxdot \hat{A})$;
  \item $\hat{A} \varoast \hat{B} : \boxdot \hat{A} \times \boxdot \hat{B} \to \boxdot (\hat{A} \times \hat{B})\text{, s. t. } \forall \hat{A}, \hat{B} \in Ob_{\mathcal{C}(\lambda)}, \hat{A} \varoast \hat{B} = [p, {\bf let \: box \:} x,y = \pi_1 p, \pi_2 p {\: \bf in \:} \langle x, y \rangle] \in Hom_{\mathcal{C}(\lambda)}(\boxdot \hat{A} \times \boxdot \hat{B}, \boxdot (\hat{A} \times \hat{B}))$.
\end{enumerate}
\end{defin}

@@

\begin{small}
\begin{prooftree}
\AxiomC{$p : \Box A \times \Box B \vdash p : \Box A \times \Box B$}
\RightLabel{$\times_E$}
\UnaryInfC{$p : \Box A \times \Box B \vdash \pi_1 p : \Box A$}
\AxiomC{$p : \Box A \times \Box B \vdash p : \Box A \times \Box B$}
\RightLabel{$\times_E$}
\UnaryInfC{$p : \Box A \times \Box B \vdash \pi_2 p : \Box B$}
\AxiomC{$x : A \vdash x : A$}
\AxiomC{$y : B \vdash y : B$}
\RightLabel{$\times_I$}
\BinaryInfC{$x : A, y : B \vdash \langle x, y \rangle : A \times B$}
\RightLabel{${\bf let}_{\Box}$}
\TrinaryInfC{$p : \Box A \times \Box B \vdash {\bf let \: box \:} x,y = \pi_1 p, \pi_2 p {\: \bf in \:} \langle x, y \rangle : \Box (A \times B)$}
\end{prooftree}
\end{small}
